---
# output:
#  bookdown::word_document2: default
params:
  title: "title"
  author: "author"
  start.date: "start.date"
  end.date: "end.date"
  dac: "DAC"
  nih.dac.action.table: "nih_dac_action_table"
  timeline.summary.table: "timeline.summary.table"
  study.summary.table: "study.summary.table"
  top.study.term.freq.table: "top.study.term.freq.table"
  study.status.table.all: "study.status.table.all"
  study.status.table.dac: "study.status.table.dac"
  pi.requests.table.overall: "pi.requests.table.overall"
  pi.requests.table.selected.time: "pi.requests.table.selected.time"
  approval.time.diff.table: "approval.time.diff.table"
  doc.type: "docx"
--- 

---
title: `r params$title`
author: `r params$author`
date: `r format(Sys.time(), '%d %B, %Y')`
---

```{r, include=FALSE}
# Loads variables to be reported
library(ggplot2)
library(plotly)
# util function
get.df.within.range <- function(df,start.date,end.date,date.col="Approved by DAC") {
  return(subset(df, as.Date(df[,date.col], format="%m/%d/%Y") >= start.date & as.Date(df[,date.col], format="%m/%d/%Y") <= end.date))
}

dac <- params$dac
nih_dac_action_table <- params$nih.dac.action.table
dac.total.dars <- nrow(nih_dac_action_table[nih_dac_action_table['DAC'] == dac,])

total.number.of.projects <- length(unique(nih_dac_action_table[nih_dac_action_table['DAC'] == dac,'Project']))
dac.row <- timeline.summary.table[timeline.summary.table['DAC'] == dac,]

num.dars.in.window <- dac.row[,'TotalRequests']
number.of.accepted <- dac.row[,'TotalApproved']
number.of.rejected <- dac.row[,'TotalRejected']
number.of.downloaded <- dac.row[,'Downloaded']
number.of.prev.download <- dac.row[,'PreviouslyDownloaded']

avg.accept <- as.numeric(dac.row[,'AvgApprovalTime'])
avg.reject <- as.numeric(dac.row[,'AvgRejectionTime'])
avg.final <- (avg.accept*number.of.accepted + avg.reject*number.of.rejected) / (number.of.accepted+number.of.rejected)
```

The `r dac` Data Access Committee (DAC) currently manages `r dac.total.dars` data access requests (DARs) for access to `r total.number.of.projects` projects in dbGaP.  

# Data Access Requests

Between `r params$start.date` and `r params$end.date` `r dac` reviewed `r num.dars.in.window` DARs. Of these, `r number.of.accepted` were accepted, `r number.of.downloaded` were downloaded, `r number.of.prev.download` had a previous version downloaded, and `r number.of.rejected` were rejected.   The average amount of time from when the Principle Investigator (PI) submited a DAR to the final decision by the DAC was `r round(avg.final,1)` days.  The average time to an accepted decision was `r round(avg.accept,1)` days, while the average time to a rejected decision was `r round(avg.reject,1)` days.  Figure \@ref(fig:compline) is a barplot comparing the `r dac` DAC to time to final decision to the average across all NIH DACs during the same time interval.  

```{r compline, fig.cap="Comparison of DAR Processing Time among all DACs", echo=FALSE, warning=FALSE ,message=FALSE, fig.align="center"}
timeline.summary.table <- timeline.summary.table %>% dplyr::mutate(ToHighlight = ifelse( DAC == params$dac, "yes", "no" ))
timeline.summary.table$AvgApprovalTime <- as.numeric(timeline.summary.table$AvgApprovalTime)
timeline.summary.table<- timeline.summary.table[order(timeline.summary.table$AvgApprovalTime, decreasing=TRUE),]
timeline.summary.table$DAC <- factor(timeline.summary.table$DAC, levels=unique(as.character(timeline.summary.table$DAC)))
p <- ggplot(data=timeline.summary.table, aes(x=DAC, y=AvgApprovalTime, fill = ToHighlight)) +
  geom_bar(stat="identity") +
  # geom_text(aes(label=paste(round(AvgApprovalTime,1),'days'))) +
  scale_y_continuous(breaks= seq(0, max(timeline.summary.table$AvgApprovalTime, na.rm = TRUE), by = 7)) +
  scale_fill_manual(values = c("yes"="cornflowerblue","no"="gray"), guide = FALSE) +
  coord_flip() +
  ylab("Average Processing Time (days)")
if (params$doc.type == "html") {
  ggplotly(p)
}
if (params$doc.type == "docx") {
  p
}
```

```{r include=FALSE}
dar.access.fig.cap <- sprintf("Data Access Requests Submitted to %s Per Month Since 2015", dac)
```


```{r fig.cap=dar.access.fig.cap, echo=FALSE, warning=FALSE ,message=FALSE, fig.align="center"}
dar.request.growth.plot <- ggplot(study.status.table.dac, aes(x=Month,y=TotalRequests)) +
  geom_line(size=2, color="cornflowerblue") +
  xlab("Year") +
  ylab("DAR Total")
rect <- data.frame(xmin=as.Date(lubridate::floor_date(as.Date(params$start.date))), xmax=as.Date(lubridate::floor_date(as.Date(params$end.date))), ymin=0, ymax=Inf)
dar.request.growth.plot <- dar.request.growth.plot + geom_rect(data=rect, aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax),
              color="transparent", fill="orange", alpha=0.3, inherit.aes = FALSE)
if (params$doc.type == "html") {
  ggplotly(dar.request.growth.plot)
}
if (params$doc.type == "docx") {
  dar.request.growth.plot
}
```


```{r fig.cap="DAR Processing Time: From PI Submission to SO Approval", echo=FALSE, warning=FALSE ,message=FALSE, fig.align="center"}
# calculates the highest whisker for all boxplots
approval.time.top.ylims <- stats::aggregate(approval.time.diff.table["ApprovalTimeFromPIToSO"],
                                           by=list(Month=approval.time.diff.table$ApprovedMonth),
                                           FUN = function(x){ boxplot.stats(as.numeric(x))$stats[5] }
                                           )
# Take the max one and use it as ylim
ylim1 <- c(0,max(approval.time.top.ylims$ApprovalTimeFromPIToSO))
#ylim1 <- boxplot.stats(as.numeric(approval.time.diff.table$ApprovalTimeFromPIToSO))$stats[c(1, 5)]


approval.time.from.PI.to.SO.plot <- ggplot(approval.time.diff.table, aes(factor(ApprovedMonth), ApprovalTimeFromPIToSO)) +
  geom_boxplot(outlier.shape = NA) + 
  coord_cartesian(ylim = ylim1*1.05) +
  scale_y_continuous(n.breaks = 8) +
  stat_summary(fun=median, colour="cornflowerblue", aes(y=ApprovalTimeFromPIToSO,group=1),
               geom="line", group=1, size=2) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  xlab("Month") +
  ylab("Time (Days)")

if (params$doc.type == "html") {
  ggplotly(approval.time.from.PI.to.SO.plot)
}
if (params$doc.type == "docx") {
  approval.time.from.PI.to.SO.plot
}
```

```{r fig.cap="DAR Processing Time: From SO Approval to DAC Approval", echo=FALSE, warning=FALSE ,message=FALSE, fig.align="center"}
approval.time.top.ylims <- stats::aggregate(approval.time.diff.table["ApprovalTimeFromSOToDAC"],
                                           by=list(Month=approval.time.diff.table$ApprovedMonth),
                                           FUN = function(x){ boxplot.stats(as.numeric(x))$stats[5] }
                                           )

ylim2 <- c(0,max(approval.time.top.ylims$ApprovalTimeFromSOToDAC))

approval.time.from.SO.to.DAC.plot <- ggplot(approval.time.diff.table, aes(factor(ApprovedMonth), ApprovalTimeFromSOToDAC)) +
  geom_boxplot(outlier.shape = NA) + 
  coord_cartesian(ylim = ylim2*1.05) +
  scale_y_continuous(n.breaks = 8) +
  stat_summary(fun=median, colour="cornflowerblue", aes(y=ApprovalTimeFromSOToDAC,group=1),
               geom="line", group=1, size=2) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  xlab("Month") +
  ylab("Time (Days)")

if (params$doc.type == "html") {
  ggplotly(approval.time.from.SO.to.DAC.plot)
}
if (params$doc.type == "docx") {
  approval.time.from.SO.to.DAC.plot
}
```

```{r fig.cap="DAR Processing Time: From PI Submission to DAC Approval", echo=FALSE, warning=FALSE ,message=FALSE, fig.align="center"}
approval.time.top.ylims <- stats::aggregate(approval.time.diff.table["ApprovalTimeFromPIToDAC"],
                                           by=list(Month=approval.time.diff.table$ApprovedMonth),
                                           FUN = function(x){ boxplot.stats(as.numeric(x))$stats[5] }
                                           )

ylim3 <- c(0,max(approval.time.top.ylims$ApprovalTimeFromPIToDAC))

approval.time.from.PI.to.DAC.plot <- ggplot(approval.time.diff.table, aes(factor(ApprovedMonth), ApprovalTimeFromPIToDAC)) + geom_boxplot(outlier.shape = NA) + 
  coord_cartesian(ylim = ylim3*1.05) +
  # scale_x_discrete(labels = format(as.Date(approval.time.diff.table$ApprovedMonth), "%m/%Y")) +
  scale_y_continuous(n.breaks = 8) +
  stat_summary(fun=median, colour="cornflowerblue", aes(y=ApprovalTimeFromPIToDAC,group=1),
               geom="line", group=1, size=2) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  xlab("Month") +
  ylab("Time (Days)")

if (params$doc.type == "html") {
  ggplotly(approval.time.from.PI.to.DAC.plot)
}
if (params$doc.type == "docx") {
  approval.time.from.PI.to.DAC.plot
}

```

```{r echo=FALSE, warning=FALSE ,message=FALSE}

nih_dac_action_table_current <- get.df.within.range(nih_dac_action_table,start.date,end.date,date.col="Submitted by PI")
#print(nih_dac_action_table_current)

pi.submitted.to.current.dac.selected.time <- unique(nih_dac_action_table_current[nih_dac_action_table_current["DAC"] == dac,]$PI)
#print(pi.submitted.to.current.dac.selected.time)

current.dac.pi.table <- subset(pi.requests.table.selected.time, PI %in% pi.submitted.to.current.dac.selected.time)
#print(current.dac.pi.table)

current.dac.pi.submit.to.more.than.one.dac <-  sum(current.dac.pi.table$DAC > 1)
current.dac.pi.average.dar <- round(mean(current.dac.pi.table$DAR),2)
current.dac.pi.average.dac <- round(mean(current.dac.pi.table$DAC),2)
current.dac.pi.average.project <- round(mean(current.dac.pi.table$Project),2)
```

Between `r start.date` and `r end.date`, `r length(pi.submitted.to.current.dac.selected.time)` PIs have submitted DAR to studies released by `r dac`. Among these PIs, `r current.dac.pi.submit.to.more.than.one.dac` have also submitted DAR to studies released by other DACs. PIs who submitted DARs to `r dac` also
submits, on average, `r current.dac.pi.average.dar` DARs to all DACs for `r current.dac.pi.average.project` projects.

# Study Released

```{r include=FALSE}
# study.summary.table.ordered <- study.summary.table[order(-study.summary.table$TotalRequest),]
# total.approved.dar <- sum(study.summary.table.ordered$TotalApproved)
# total.dar.downloaded <- sum(study.summary.table.ordered$TotalDownload)
# download.percentage <- round(total.dar.downloaded*100/total.approved.dar,2)
# phs.popular.all <- study.summary.table.ordered[1,]['StudyAccesion']
# phs.popular.all.study.name <- study.summary.table.ordered[1,]['Study Name']
# phs.popular.all.dac <- study.summary.table.ordered[1,]['DAC']
# phs.popular.all.total.request <- study.summary.table.ordered[1,]['TotalRequest']

dac.study.summary.table <- study.summary.table[study.summary.table['DAC'] == dac,]
dac.study.summary.table <- get.df.within.range(dac.study.summary.table,params$start.date,params$end.date,"Study Release Date")
new.studies.released <- nrow(dac.study.summary.table)

new.studies.dar <- sum(dac.study.summary.table$TotalRequest, na.rm = TRUE)
dac.study.summary.table.ordered <- dac.study.summary.table[order(-dac.study.summary.table$TotalRequest),]
if (nrow(dac.study.summary.table.ordered) == 0) {
  phs.popular.this.dac <- NA
  phs.popular.this.dac.study.name <- NA
  phs.popular.this.dac.total.request <- NA
} else {
  phs.popular.this.dac <- dac.study.summary.table.ordered[1,]['StudyAccesion']
  phs.popular.this.dac.study.name <- dac.study.summary.table.ordered[1,]['Study Name']
  phs.popular.this.dac.total.request <- dac.study.summary.table.ordered[1,]['TotalRequest']
}

```

During this reporting period, `r new.studies.released` new studies were released by `r dac` through dbGaP.  A total of `r new.studies.dar` DARs were made for these newly released studies. Study `r phs.popular.this.dac` (`r phs.popular.this.dac.study.name`) was the most requested dataset from the `r dac` DAC with `r phs.popular.this.dac.total.request` requests.  The following word-cloud provides a graphical summary of the combined research use statements (RUS) from all DARs requesting this study. The larger the word appears in the wordcloud, the more frequent the word in the set of RUS.

```{r include=FALSE}
study.released.fig.cap <- sprintf("Number of Study Released by %s Per Month Since 2015", dac)
word.cloud.fig.cap <- sprintf("Word cloud of Research Statements for Study %s (%s)",phs.popular.this.dac,phs.popular.this.dac.study.name)
```

```{r fig.cap=word.cloud.fig.cap, echo=FALSE, warning=FALSE ,message=FALSE, fig.align="center"}
library(wordcloud)
library(RColorBrewer)
if (!is.na(top.study.term.freq.table)) {
  wordcloud(words = top.study.term.freq.table$word, freq = top.study.term.freq.table$freq, min.freq = 1, max.words=200,
            random.order=FALSE, rot.per=0.1, colors=RColorBrewer::brewer.pal(8, "Dark2"))
}
```

```{r include=FALSE}
dac.study.summary.table.ordered.by.citation <- dac.study.summary.table[order(-(dac.study.summary.table$gs_citation_count)),]
if (nrow(dac.study.summary.table.ordered.by.citation) > 0) {
  total.citations <- sum(dac.study.summary.table.ordered.by.citation$`gs_citation_count`, na.rm=TRUE)
  most.cited.study <- dac.study.summary.table.ordered.by.citation[1,]
  most.cited.study.id <- most.cited.study$`StudyAccesion`
  most.cited.study.name <- most.cited.study$`Study Name`
  most.cited.study.count <- most.cited.study$`gs_citation_count`
} else {
  total.citations <- 0
  most.cited.study.id <- NA
  most.cited.study.name <- NA
  most.cited.study.count <- NA
}

```

According to Google Scholar, the most cited study during this time period is `r most.cited.study.id` (`r most.cited.study.name`) with `r most.cited.study.count` citations.

```{r include=FALSE}
most.cited.fig.cap <- sprintf("Top 5 Studies Managed by %s with Most Google Scholar Citations Between %s and %s", dac,params$start.date,params$end.date)

if (nrow(dac.study.summary.table.ordered.by.citation) > 5) {
  most.cited.studies <- dac.study.summary.table.ordered.by.citation[1:5,]
  citation.count <- as.vector(most.cited.studies$`gs_citation_count`)
  study.ids <-  as.vector(most.cited.studies$`StudyAccesion`)
  labels <- as.vector(most.cited.studies$`Study Name`)
  other.total <- total.citations - sum(citation.count)
  if (other.total > 0) {
      citation.count <- c(citation.count,other.total)
      labels <- c(labels,"others")
      study.ids <- c(study.ids,"others")
  }
  
  cited.df <- data.frame(StudyID = study.ids, 
                         StudyName = labels,
                         Citations = citation.count
                         )
  
  # cited.df <- cited.df[order(cited.df$Citations,decreasing = TRUE),]
  cited.df$StudyID <- factor(cited.df$StudyID,
                                       levels=unique(as.character(cited.df$StudyID)))
}
```

```{r fig.cap=most.cited.fig.cap, echo=FALSE, fig.align="center"}
  show.legend <- doc.type=="docx"
  cited.plot <- ggplot(data=cited.df, aes(x = StudyID, y = Citations, fill = StudyName)) +
    geom_bar(stat="identity",show.legend = show.legend) +
    geom_text(aes(label=Citations), vjust=-0.3, size=3.5) +
    scale_y_continuous(n.breaks = 8) +
    scale_x_discrete(guide = guide_axis(n.dodge=2)) +
    xlab("Study IDs") +
    ylab("Citations")

  if (params$doc.type == "html") {
    cited.plot <- cited.plot + theme(legend.position='none')
    ggplotly(cited.plot)
  }
  if (params$doc.type == "docx") {
    cited.plot <- cited.plot +
      theme(legend.position = "bottom") +
      guides(fill = guide_legend(nrow = 6,title.position="top", title.hjust = 0.5))
    cited.plot 
  }
```


```{r fig.cap=study.released.fig.cap, echo=FALSE, warning=FALSE ,message=FALSE, fig.align="center"}
study.released.growth.plot <- ggplot(study.status.table.dac, aes(x=Month,y=StudiesReleased)) +
  geom_line(size=2, color="cornflowerblue") +
  xlab("Year") +
  ylab("Studies Released")
rect <- data.frame(xmin=as.Date(lubridate::floor_date(as.Date(params$start.date))), xmax=as.Date(lubridate::floor_date(as.Date(params$end.date))), ymin=0, ymax=Inf)
study.released.growth.plot <- study.released.growth.plot + geom_rect(data=rect, aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax),color="transparent", fill="orange", alpha=0.3, inherit.aes = FALSE)

if (params$doc.type == "html") {
  ggplotly(study.released.growth.plot)
}
if (params$doc.type == "docx") {
  study.released.growth.plot
}
```


# Other Comparisons


```{r include=FALSE}
dual.y.fig.caption <- sprintf("Comparison of Cummulative Requests Received and Studies Released by %s", dac)
```


```{r fig.cap=dual.y.fig.caption, echo=FALSE, warning=FALSE, message=FALSE, fig.align="center"}
# A few constants
coeff <- 100
coeff <- max(study.status.table.dac$TotalRequestsCummulative)/max(study.status.table.dac$StudiesReleasedCummulative)

comparison.plot <- ggplot(study.status.table.dac, aes(x=Month)) +
  geom_line(aes(y=StudiesReleasedCummulative), size=2, color="coral3") +
  geom_line(aes(y=TotalRequestsCummulative/coeff), size=2, color="cornflowerblue") +
  scale_y_continuous(
    # Features of the first axis
    name = "Studies Released\n",
    breaks = scales::pretty_breaks(n = 8),
    # Add a second axis and specify its features
    sec.axis = sec_axis(~.*coeff, name="Requests Received\n", breaks = scales::pretty_breaks(n = 8))
  ) +
  theme(
    axis.title.y = element_text(color = "coral3", size=13),
    axis.title.y.right = element_text(color = "cornflowerblue", size=13),
    axis.ticks.y.left = element_line("coral3"),
    axis.ticks.y.right = element_line("cornflowerblue"),
    axis.text.y.left = element_text(color = "coral3", size=13),
    axis.text.y.right = element_text(color = "cornflowerblue", size=13),
  ) +
  xlab("Year")

rect3 <- data.frame(xmin=as.Date(lubridate::floor_date(as.Date(params$start.date))), xmax=as.Date(lubridate::floor_date(as.Date(params$end.date))), ymin=0, ymax=Inf)

comparison.plot <- comparison.plot + geom_rect(data=rect3, aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax),color="transparent", fill="orange", alpha=0.3, inherit.aes = FALSE)

if (params$doc.type == "html") {
  ggplotly(comparison.plot)
}
if (params$doc.type == "docx") {
  comparison.plot
}

```

```{r include=FALSE}
growth.compare.caption <- sprintf("Comparison of DAR Growth and Studies Released Growth for %s",dac)
```

```{r fig.cap=growth.compare.caption, echo=FALSE, warning=FALSE, message=FALSE, fig.align="center"}
study.status.table.dac.plot <- ggplot(study.status.table.dac, aes(x=Month)) +
  geom_line(aes(y=TotalRequestsGrowth, colour="Total Requests Growth"), size=2) +
  geom_line(aes(y=StudiesReleasedGrowth, colour="Studies Released Growth"), size=2) +
  scale_color_manual("",
                     values = c("Total Requests Growth"="cornflowerblue","Studies Released Growth"="coral3")) +
  theme(legend.position = "bottom") +
  scale_x_date(date_breaks = "6 month", date_labels = '%m/%Y') +
  scale_y_continuous(n.breaks = 8, labels = function(x) paste0(x*100, "%")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  xlab("Year") +
  ylab("Growth (%)")
  
rect2 <- data.frame(xmin=as.Date(lubridate::floor_date(as.Date(params$start.date))), xmax=as.Date(lubridate::floor_date(as.Date(params$end.date))), ymin=0, ymax=Inf)

study.status.table.dac.plot <- study.status.table.dac.plot + geom_rect(data=rect2, aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax),color="transparent", fill="orange", alpha=0.3, inherit.aes = FALSE)

if (params$doc.type == "html") {
  ggplotly(study.status.table.dac.plot)
}
if (params$doc.type == "docx") {
  study.status.table.dac.plot
}
```

---

This report was prepared using the DACReportingTool package for R, build `r packageVersion("DACReportingTool")` by Mr. Hoyin Chu and Dr. Christopher Steven Marcum. 
